---
title: "Lab 3 Report"
author: "Logan Hillers"
date: "`r format(Sys.Date())`"
output: github_document
---

Instructions for this lab report can be found on [Your assignment section of Lab Assignment 3 Introduction to ggplot2](https://biol275-msum.github.io/introduction-to-ggplot2.html#your-assignment) on the lab website.

## Fireflies

> A. Insert an R code chunk and create a graph depicting the frequency distribution of the 35 mass measurements. It should have legible text and appropriate axis labels.

First the tidyverse packages were ran
```{r}
library(tidyverse)
```
Next I uploaded the firefly data from a CSV document
```{r}
firefly_mass <- read_csv("https://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter02/chap02q19FireflySpermatophoreMass.csv")

firefly_mass
```
I then used the packages and the uploaded data to create a histogram
```{r}
ggplot(data = firefly_mass) +
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.01,
                 boundary = 0, closed = "left", 
                 fill = "#C5351B", color = "black") +
  labs(x = "Spermatophore Mass", y = "Frequency (number of fireflies)") +
  scale_y_continuous(breaks = seq(0, 30, 5), limits = c(0, 12), 
                     expand = expansion(mult = 0)) +
  scale_x_continuous(breaks = seq(0, 600, 100)) +
  theme_classic() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1))
  )
```


> B. What type of graph did you choose in part (A)? Why?

I chose a histogram for part (A) because I wanted to visualize the frequency of each spermatophore mass that occurred in the sample, and histograms are great at displaying numerical data that are a range of values.

> C. Describe the shape of the frequency distribution. What are its main features?

The shape of the frequency distribution appears to have two main peaks, making it bimodal.

> D. What term would be used to describe the largest measurement in the frequency distribution? You can refer to the lecture slides, your notes, or the textbook.

The mode is the largest measurement in the frequency distribution.

## Bird orders

I read the data using the code provided by the instructor. The dataset is from the auk package and was originally named ebird_taxonomy. I converted it to a tibble, removed non-species taxa, and named the new dataset birds.


> E. How many bird species are in the new birds dataset? How did you find out?

There are 10,721 bird species in the new bird dataset, which I found by loading the data in my console by running the code below.

```{r}
count(birds)
```


> H. How many orders are there? You can get this right by counting on the graph, but you will only get full points if you use some code to figure it out.

I counted how many orders there were by running the code below. In total, there were 41 orders.

```{r}
count(birds, order)
```


> F. Insert an R code chunk and create a graph depicting the distribution of orders in the birds dataset. Sort the orders with the most frequent on the left. It should have legible text and appropriate axis labels.

I read the dataset from the auk package. I changed it to a tibble, removed non-species taxa, and named the new dataset birds. I did this after the tidyverse packages were installed.

```{r}
library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B", 
           width = .8) +
  labs(x = "Order", y = "Frequency (number of species)") +
  scale_y_continuous(limits = c(0, 6500), expand = expansion(mult = 0)) +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1)),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_blank()
  )
```

> G. What kind of graph did you create? Why?

I created a bar graph because I was comparing the number of species in each order, which is a discrete numerical value, and has nothing to do with frequency. Each bar represents each order, so the bar sizes can be compared easily by the eye and the one can infer which orders are larger than the others.

## Links to peer review

I reviewed Faith's lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-CForEffort.git

Faith reviewed my lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-LoganHillers.git

## Session Info

Here is my session information:

```{r}
sessioninfo::session_info()
```
